---
title: "Simulation Exercise"
author: "Tejus"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

## Title :  

### Overview : 
- Investigating the exponential distribution 
- In R we can use the distribution using `rexp(n, lambda)`, where `lambda` is the rate parameter.
- Assuming that the population & the randomness governing our sample can be given by densities and mass functions. We work with characteristics of these densities & mass functions instead of the whole function.  

**Key characteristics about the distribution :**

- Mean of the exponential distribution *1 / lambda* [Characterization of it’s center]
- Standard deviation is also *1 / lambda* [Characterization of spread of data.]

*Note : Setting lambda = 2 for all the simulations & averages of 40 exponentials will be investigated by simulating 1000 times*


```{r Constants, echo=FALSE}
lambda <- 0.2
n <- 40
n_sim <- 1000

# Random numbers from the distribution
ExpDataPts <- matrix(rexp(n_sim, lambda)) # Exp dist
ExpDataPts40 <- apply( matrix(rexp(n_sim*n), n_sim),
                        1, mean)
NormalDataPts <- matrix(rnorm(n_sim, 1/lambda)) # Normal dist.

```

The following plot shows the Normally distributed points along with our Exponentially distributed points both with the same mean `r 1/lambda`. Black horizontal line is for the mean of those distributions.  

```{r ExpDist, echo=FALSE}
# Line plot for both
Y <- c(sort(ExpDataPts), sort(NormalDataPts))
X <- c(rep(1:n_sim,2))

PlotDf <- data.frame(Index = X, "Data_Points" = Y, Distribution = c(rep("Exponential Distribution", n_sim), rep("Normal Distribution", n_sim)))


# Plotting
g <- ggplot(data = PlotDf, aes(Index, Data_Points))
g <- g + geom_point(aes(color = Distribution)) + geom_hline(yintercept = 1/lambda, size = 1.5)
print(g)

```

- From the plot you can see that unlike Normal Distribution, data points are not symmetrical around mean. As per the name `Exponential Distribution` the data does look like it's growing exponentially.  

**Histogram of an typical exponential curve would look something like this**  

```{r Histogram, echo=FALSE}

g <- ggplot(data = PlotDf, aes(x = Data_Points)) + geom_histogram(aes(fill = Distribution), binwidth = 1)  + geom_vline(xintercept = 1/lambda, size = 1.5)

print(g) 

```

Again you can see in this histogram that the `Exponential` distribution compared to `Normal` distribution is centered around the same mean (black vertical line) but not symmetric around that.

### The sample mean and the theoretical mean of the distribution.

- The black line above is the population mean. Since we have simulated these points from the mean `r 1/lambda` we know the absolute value of it. However in real life experiments we want to know what that value is.  
- Why ? Because as I said above that mean (true mean) of a distribution is a very good characteristic to estimate so that we get an idea about the distribution; to make inferences about the population.  

**Sample Mean**

- Sample Mean is an estimate of **Population Mean** `(Theoretical Mean)`.  
- After doing some experiments & collecting data the mean of that data is our **Sample Mean** which is the estimate for our population mean.  


```{r SampleMean, echo=FALSE}

PlotDf <- data.frame(Data_Points = ExpDataPts, Index = 1:n_sim)

Stat <- data.frame(x = c(1/lambda), Line = c("Theoretical Mean"))

g <- ggplot(data=PlotDf, aes(x = Data_Points)) + 
            geom_histogram(aes(y =..density..), binwidth = 1, fill="white", color = "black") + 
            geom_density(size = 1.5, color="black", fill = "#FF6666", alpha = 0.2) + 
            geom_vline(aes(xintercept = x, linetype = Line),data=Stat, color = "BLUE", size = 1.5, show.legend = TRUE) + 
            labs(title = "Theoretical Mean amongst Sample Mean") + 
            labs(x= "Exponential Data Point Values", y = "Density") 

print(g)

```

Clearly the peak (sample mean) is close to the *Population mean* but not exactly the same. It will approach the population mean as we increase the sample size. 

### The sample variance and the theoretical variance of the distribution.

- Sample Variance is an estimate of **Population Variance**.  
- After doing some experiments & collecting data the variance of that data is our **Sample variance** which is the estimate for our population variance.
- Since it's also a random variable, it must be centered around true population variance
`r 1/lambda^2`


```{r SampleVar, echo=FALSE, fig.cap="Distribution of 1000 variances for (40 values each) "}

Var_ExpDataPts40 <- apply( matrix(rexp(n_sim*n), n_sim),
                        1, var)
PlotDf <- data.frame(Variance = Var_ExpDataPts40, Index = 1:n_sim)
Stat <- data.frame(x = c((1/lambda^2)/n,1/lambda^2), Line = c("Sample Varaince","Theoretical Variance"))

g <- ggplot(data=PlotDf, aes(x = Variance)) + 
            geom_histogram(aes(y =..density..), binwidth = 1, fill="white", color = "black") + 
            geom_vline(aes(xintercept = x, linetype = Line),data=Stat, color = "BLUE", size = 1.5, show.legend = TRUE) + 
            labs(title = "Theoretical variance amongst Sample variance") + 
            labs(x= "Variance of Exponential Data Values", y = "Density") +
            geom_density(size = 1.5, color="black", fill = "#FF6666", alpha = 0.2) 
            
print(g)

```

Sample variance is an estimate of population variance hence as we have learned that Population variance (`r 1/lambda`) / sample size (`r n`) gives us our **sample variance** (`r (1/lambda)/n`)

It is clearly not centered at population variance because in this case it was just 40 values each. But as we increase the sample size it will approach the true **Population Variance**.



### The sample variance and the theoretical variance of the distribution.



```{r CLT, echo=FALSE}
Mean_ExpDataPts40 <- apply( matrix(rexp(n_sim*n), n_sim),
                        1, mean)

PlotDf <- data.frame(Mean = c(ExpDataPts, Mean_ExpDataPts40) , Index = rep(1:n_sim, 2), "Factor" = c(rep("Avg", n_sim), rep("Avg40", n_sim)) )
PlotDf$Factor <- as.factor(PlotDf$Factor)

Stat <- data.frame(x = c(1/lambda), Line = c("Theoretical Mean"))

g <- ggplot(data=PlotDf, aes(x = Mean)) + 
            geom_histogram(aes(y =..density..), binwidth = 0.35, fill="white", color = "black") +  
            labs(title = "Law of large numbers") + 
            labs(x= "Density", y = "Value Exponential Data Values") +
            geom_density(size = 1.5, color="black", fill = "#FF6666", alpha = 0.2) +
            facet_grid(.~Factor)
            
print(g)


```


In the above plot