---
title: "Simulation Exercise"
author: "Tejus"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

### Overview : 
- Investigating the exponential distribution 
- In R we can use the distribution using `rexp(n, lambda)`, where `lambda` is the rate parameter.
- Assuming that the population & the randomness governing our sample can be given by densities and mass functions. We work with characteristics of these densities & mass functions instead of the whole function.  

**Key characteristics about the distribution :**

- Mean of the exponential distribution *1 / lambda* [Characterization of it’s center]
- Standard deviation is also *1 / lambda* [Characterization of spread of data.]

*Note : Setting lambda = 2 for all the simulations & averages of 40 exponentials will be investigated by simulating 1000 times*


```{r Constants, echo=FALSE}
lambda <- 0.2
n <- 40
n_sim <- 1000

# Random numbers from the distribution
ExpDataPts <- rexp(n_sim, lambda) # Exp dist
ExpDataPts40 <- apply( matrix(rexp(n_sim*n, lambda), n_sim),
                        1, mean)
NormalDataPts <- rnorm(n_sim, 1/lambda) # Normal dist.
NormalDataPts40 <- apply( matrix(rnorm(n_sim*n, 1/lambda), n_sim),
                        1, mean)

```

The following plot shows the Normally distributed points along with our Exponentially distributed points both with the same mean `r 1/lambda`. Black horizontal line is for the mean of those distributions.  

```{r ExpDist, echo=FALSE}
# Line plot for both
Y <- c(sort(ExpDataPts), sort(ExpDataPts40), sort(NormalDataPts),  sort(NormalDataPts40))
X <- c(rep(1:n_sim,4))

Stat <- data.frame(x = c(1/lambda), Line = c("Theoretical Mean"))

PlotDf <- data.frame(Index = X, "Data_Points" = Y, Distribution = c(rep("Exponential Distribution", n_sim*2), rep("Normal Distribution", n_sim*2)), Avg = c(rep("Mean",n_sim), rep("Mean40", n_sim), rep("Mean",n_sim), rep("Mean40", n_sim)))

Pdf <- subset(PlotDf, Avg == "Mean" )

# Plotting
g <- ggplot(data = Pdf, aes(Index, Data_Points))
g <- g + geom_point(aes(color = Distribution)) + geom_hline(aes(yintercept = x, linetype = Line), data=Stat, size = 1.5) + facet_grid(.~Avg)
print(g)

```

- From the plot you can see that unlike Normal Distribution, data points are not symmetrical around mean. As per the name `Exponential Distribution` the data does look like it's growing exponentially.  

**Histogram of an typical exponential curve would look something like this**  

```{r Histogram, echo=FALSE}

Stat <- data.frame(x = c(1/lambda), Line = c("Theoretical Mean"))
g <- ggplot(data = PlotDf, aes(x = Data_Points)) + geom_histogram(aes(fill = Distribution), binwidth = 1)  + geom_vline(aes(xintercept = x, linetype = Line), data=Stat, size = 1.5)

print(g) 

```

Again you can see in this histogram that the `Exponential` distribution compared to `Normal` distribution it is not symmetric around mean (black vertical line), but centered at the mean.

### The sample mean and the theoretical mean of the distribution.

- The black line above is the population mean. Since we have simulated these points from the mean `r 1/lambda` we know the absolute value of it. However in real life experiments we want to know what that value is.  
- Why ? Because as I said above that mean (true mean) of a distribution is a very good characteristic to estimate so that we get an idea about the distribution; to make inferences about the population.  

**Sample Mean**

- Sample Mean is an estimate of **Population Mean** `(Theoretical Mean)`.  
- After doing some experiments & collecting data, we take the mean of that data; that is our **Sample Mean** `the estimate for our population mean.`  


```{r SampleMean, echo=FALSE, fig.cap="Distribution of 1000 simulations of 1.) Exponential Values & 2.) Mean of 40 Exponential Values"}

Pdf <- subset(PlotDf, Distribution == "Exponential Distribution" )

Stat <- data.frame(x = c(1/lambda), Line = c("Theoretical Mean"))

g <- ggplot(data=Pdf, aes(x = Data_Points)) + 
            geom_histogram(aes(y =..density..), binwidth = 1, fill="white", color = "black") + 
            geom_density(size = 1.0, color="black", fill = "#FF6666", alpha = 0.2) + 
            geom_vline(aes(xintercept = x, linetype = Line),data=Stat, color = "BLUE", size = 1.0, show.legend = TRUE) + 
            labs(title = "Theoretical Mean amongst Sample Mean") + 
            labs(x= "Exponential Data Point Values", y = "Density") + facet_grid(.~Avg)

print(g)

```

Clearly the peak (sample mean) is close to the *Population mean* but not exactly the same in the first plot. It approaches the population mean as we increase the sample size (Take each value as mean of 40 such values; Inturn `r n_sim*n` total samples).  

### The sample variance and the theoretical variance of the distribution.

- Sample Variance is an estimate of **Population Variance**.  
- After doing some experiments & collecting data the variance of that data is our **Sample variance** which is the estimate for our population variance.
- Since it's also a random variable, it must be centered around true population variance
`r 1/lambda^2`


```{r SampleVar, echo=FALSE, fig.cap="Distribution of variance from 1000 simulations for mean(40 values each) "}

Var_ExpDataPts40 <- apply( matrix(rexp(n_sim*n), n_sim),
                        1, var)

PlotDf <- data.frame(Variance = Var_ExpDataPts40, Index = 1:n_sim)
Stat <- data.frame(x = c((1/lambda^2)/n,1/lambda^2), Line = c("Sample Varaince","Theoretical Variance"))

g <- ggplot(data=PlotDf, aes(x = Variance)) + 
            geom_histogram(aes(y =..density..), binwidth = 1, fill="white", color = "black") + 
            geom_vline(aes(xintercept = x, linetype = Line),data=Stat, color = "BLUE", size = 1.5, show.legend = TRUE) + 
            labs(title = "Theoretical variance amongst sample variance") + 
            labs(x= "Variance of Exponential Data Values", y = "Density") +
            geom_density(size = 1.5, color="black", fill = "#FF6666", alpha = 0.2) 
print(g)

```

Sample variance is an estimate of population variance hence as we have learned that Population variance (`r 1/lambda`) / sample size (`r n`) gives us our **sample variance** (`r (1/lambda)/n`)

It is clearly not centered at population variance because in this case it was just 40 values each. But as we increase the sample size it will approach the true **Population Variance**.

### Distributions of characteristic values


```{r CLT, echo=FALSE, fig.cap="Distribution of mean & variance from 1000 simulations of 40 values each follows a Normal distribution centered at what it's trying to estimate (Population Mean / Population Variance)"}

PlotDf <- data.frame(Mean = c(ExpDataPts40, Var_ExpDataPts40) , Index = rep(1:n_sim, 2), "Factor" = c(rep("Mean", n_sim), rep("Variance", n_sim)) )
PlotDf$Factor <- as.factor(PlotDf$Factor)

g <- ggplot(data=PlotDf, aes(x = Mean)) + 
            geom_histogram(aes(y =..density..), binwidth = 0.35, fill="white", color = "black") +  
            labs(title = "Law of large numbers/ Central Limit Thm") + 
            labs(x= "Density", y = "Value Exponential Data Values") +
            geom_density(size = 1.5, color="black", fill = "#FF6666", alpha = 0.2) +
            facet_grid(.~Factor)
            
print(g)


```


In this plot we can see that, when we take the distribution of average from 40 random exponential variables we get a `Normal distribution`.  

