---
title: "Simulation Exercise"
author: "Tejus"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

## Title :  

### Overview : 
- Investigating the exponential distribution & comparing it with the Central Limit Theorem.
- In R we can use the distribution using `rexp(n, lambda)`, where `lambda` is the rate parameter.
- Assuming that the population & the randomness governing our sample can be given by densities and mass functions. We work with characteristics of these densities & mass functions instead of the whole function.  

**Key characteristics about the distribution :**

- Mean of the exponential distribution *1 / lambda* [Characterization of it’s center]
- Standard deviation is also *1 / lambda* [Characterization of spread of data.]

*Note : Setting lambda = 2 for all the simulations & averages of 40 exponentials will be investigated by simulating 1000 times*


```{r Constants, echo=FALSE}
lambda <- 2
n <- 40
n_sim <- 1000

# Random numbers from the distribution
ExpDataPts <- matrix(rexp(n_sim, lambda)) # Exp dist
NormalDataPts <- matrix(rnorm(n_sim, 1/lambda)) # Normal dist.

```

The following plot shows the Normally distributed points along with our Exponentially distributed points both with the same mean `r 1/lambda`. Black horizontal line is for the mean of those distributions.  

```{r ExpDist, echo=FALSE}
# Line plot for both
Y <- c(sort(ExpDataPts), sort(NormalDataPts))
X <- c(rep(1:n_sim,2))

PlotDf <- data.frame(Index = X, "Data_Points" = Y, Distribution = c(rep("Exponential Distribution", n_sim), rep("Normal Distribution", n_sim)))


# Plotting
g <- ggplot(data = PlotDf, aes(Index, Data_Points))
g <- g + geom_point(aes(color = Distribution)) + geom_hline(yintercept = 1/lambda, size = 1.5)
print(g)

```

- From the plot you can see that unlike Normal Distribution, data points are not symmetrical around mean. As per the name `Exponential Distribution` the data does look like it's growing exponentially.  

**Histogram of an typical exponential curve would look something like this**  

```{r Histogram, echo=FALSE}

g <- ggplot(data = PlotDf, aes(x = Data_Points)) + geom_histogram(aes(fill = Distribution), binwidth = 0.35)  + geom_vline(xintercept = 1/lambda, size = 1.5)

print(g) 

```

Again you can see in this histogram that the `Exponential` distribution compared to `Normal` distribution is centered around the same mean (black vertical line) but not symmetric around that.

### The sample mean and the theoretical mean of the distribution.

- The black line above is the population mean. Since we have simulated these points from the mean `r 1/lambda` we know the absolute value of it. However in real life experiments we want to know what that value is.  
- Why ? Because as I said above that mean (true mean) of a distribution is a very good characteristic to estimate so that we get an idea about the distribution; to make inferences about the population.  

**Sample Mean**

- Sample Mean is an estimate of **Population Mean** `(Theoretical Mean)`.  
- After doing some experiments & collecting data the mean of that data is our **Sample Mean** which is the estimate for our population mean.

#### Sample Mean versus Theoretical Mean: 

```{r SampleMean, echo=FALSE}

PlotDf <- data.frame(Data_Points = ExpDataPts, Index = 1:n_sim)

Stat <- data.frame(x = c(1/lambda), Line = c("Theoretical Mean"))

g <- ggplot(data=PlotDf, aes(x = Data_Points)) + 
            geom_histogram(aes(y =..density..), binwidth = 0.13, fill="white", color = "black") + 
            geom_density(size = 1.5, color="black", fill = "#FF6666", alpha = 0.2) + 
            geom_vline(aes(xintercept = x, linetype = Line),data=Stat, color = "BLUE", size = 1.5, show.legend = TRUE) + 
            labs(title = "Theoretical Mean amongst Sample Mean") + 
            labs(x= "Density", y = "Exponential Data Point Values") 

print(g)

```

Clearly the peak (sample mean) is close to the *Population mean* but not exactly the same. It will approach the population mean as we increase the sample size. 


```{r CentralLimitThm}

```
## Simulations



```{r pressure, echo=FALSE}

```



## Sample Variance versus Theoretical Variance:

## Distribution: